The ``Dark Side'' of Phylogenetic Comparative Methods

### 2014-08-31 Manuscript/dark.tex
Added citations

More recently however, the PCM community has moved towards mostly implementing their methods in R (cite), and code sharing has become almost ubiquitous. 
The number of R packages for PCMs has increased exponentially since 2005 when APE (cite) was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
This has led to an increasing number of methods-users jumping straight to the implementation of a method, without really carefully trying to understand what the method is doing, what its assumptions might be, or what the results they extract really mean in a biological context.
This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
The APE book (cite) for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation - cite).\\
We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology. It is extremely frustrating reading/reviewing papers where the authors have clearly not carefully read or understood the primary literature on the method they are using. 
Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available. This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better (\citealp{freckleton2009seven}, for example, mentions R as a way of improving things). %needs a bit of finessing

>>>>

More recently however, the PCM community has moved towards mostly implementing their methods in R \citep{R-Core-Team:2014aa}, and code sharing has become almost ubiquitous. 
The number of R packages for PCMs has increased exponentially since 2005 when APE \citep{Paradis:2004aa} was released, and has increased particularly sharply since 2008 (Figure \ref{PCMRpackages}).
Simultaneously, more and more methods-users are able to use R, thus when a new method is published they can now take an R package ``off the shelf'' and use it to run the method immediately.
This has led to an increasing number of methods-users jumping straight to the implementation of a method, without really carefully trying to understand what the method is doing, what its assumptions might be, or what the results they extract really mean in a biological context.
This problem is exacerbated by the fact that manuals, vignettes and help files for R packages rarely mention the assumptions of the method, or how to test model fit. 
The APE book \citep{paradis2011analysis} for example, provides no guidance on assumption testing in its chapter on phylogenetic independent contrasts, even though the methods needed to do this are well-established and very easily implemented in R (for a counter example see CAIC and caper documentation; \citep{purvis1995comparative,Orme:2013aa}).\\
We believe this issue is probably one of the biggest currently facing phylogenetic comparative biology. It is extremely frustrating reading/reviewing papers where the authors have clearly not carefully read or understood the primary literature on the method they are using. 
Additionally, many methods-users seem to try to retrofit a question so they can use the newest shiniest methods available. This is all driven by our ``publish or perish'' culture, and the pressure to produce as many papers as possible out of a dataset.
The truly sad thing about this state of affairs is that by making science more open and sharing code (something we are all passionate about and strongly support) we seem to have made the situation worse rather than better (\citealp{freckleton2009seven}, for example, mentions R as a way of improving things). %needs a bit of finessing
